Unknown

### 2024-03-06 ms/ms.tex
working on materials and methods

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\addtokomafont{disposition}{\rmfamily}
\usepackage{siunitx}
\newcommand{\gsc}{$\g_text{sc}$}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Unknown},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{notes}{%
\subsection{Notes}\label{notes}}

@@

\hypertarget{main}{%
\subsection{Main}\label{main}}

@@

We distinguished among these hypotheses by comparing AA among wild
tomato species from different native light habitats, grown under
simulated sun and shade light treatments, and measured under contrasting
light intensity (Figure of hypotheses and predictions). We measured AA
on 600 individual plants from 30 accessions (average of 10 replicates
per light treatment) using a recently developed method
{[}@triplett\_amphistomy\_2024{]}. With this method, we directly compare
the photosynthetic rate of an untreated amphistomatous leaf to that of
the same leaf with gas exchange blocked through the adaxial (upper)
surface by transparent plastic, which we refer to as `pseudohypostomy'.
To compare amphi- and pseudohypostomatous leaves at identical whole-leaf
\(g_\text{sc}\), we measure \(A\) over a range of \(g_\text{sc}\),
inducing stomatal opening and closure by modulating humidity (see
Materials and Methods for further details).

@@

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

@@

\hypertarget{accessions}{%
\subsubsection{Accessions}\label{accessions}}

We compared AA among 30 ecologically diverse accessions of wild tomato,
including representatives of all described species of \emph{Solanum}
sect. \emph{Lycopersicon} and sect. \emph{Lycopersicoides}
{[}@peralta\_taxonomy\_2008{]} and the cultivated tomato \emph{S.
lycopersicum} var. \emph{lycopersicum} (tab:accessions). Due to
constraints on growth space and time, we spread out measurements over 80
weeks from May 1, 2022 to October 31, 2023. Replicates within accession
were evenly spread out over this period to prevent confounding of
temporal variation in growth conditions with accession. {[}anything else
to say here? maybe explain accession selection and phylogeny?{]}

@@

\end{table}

\hypertarget{plant-growth-conditions}{%
\subsubsection{Plant growth conditions}\label{plant-growth-conditions}}

@@

\hypertarget{germination-and-seedling-stage}{%
\paragraph{Germination and seedling
stage}\label{germination-and-seedling-stage}}

@@

\hypertarget{light-treatments}{%
\paragraph{Light treatments}\label{light-treatments}}

@@

We assigned sun and shade treatment to lower and upper racks of a
\(\qty{1.22}{\meter} \times \qty{2.44}{\meter}\) shelving unit in a
climate-controlled growth room. We assigned the sun treatment to the
lower rack to limit diffuse light from reaching the shade treatment. The
average daytime \(\mathrm{PPFD}\) was
\(\qty{800}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\) and
\(\qty{125}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\) for sun
and shade treatments, respectively. To isolate the effect of light
intensity from quality, we used the same LED model with the the same
spectrum (Fluence SPYDR 2i, Austin, Texas, USAS), but dimmed the lights
in the shade treatment. To maintain homogeneous environmental conditions
other than light, we mixed air within the growth room using an air
circulator (Vornado 693DC, Andover, Kansas, USA) and within racks using
a miniature oscillating air circulator (Vornado Atom 1, Andover, Kansas,
USA). Despite these efforts, the air in the sun treatment was on average
\(\qty{1}{\degreeCelsius}\) warmer and the average RH was consequently 5
lower. However, because of evaporative cooling, the leaves in the sun
treatment were only \(\qty{1}{\degreeCelsius}\) on average (\(n = 600\)
leaves).

\hypertarget{leaf-trait-measurements}{%
\subsubsection{Leaf trait measurements}\label{leaf-trait-measurements}}

We selected a fully expanded, unshaded leaf at least six leaves above
the cotyledons during early vegetative growth. This typically meant that
plants had grown in light treatments for \(\approx\) 4 weeks, ensuring
they had time to sense and respond developmentally to the light
intensity of the treatment rather than the seedling conditions
{[}@schoch\_dependence\_1980{]}. Shade plants grew slower than sun
plants, hence leaves at the same developmental stage were measured on
chronologically older plants in the shade treatment. In some sun plants,
we had to use leaves higher on the stem because short internodes made
lower leaves inaccessible with the gas exchange equipment. We measured
terminal leaflets in 70.0\% of cases, but used the lateral leaflet
closest to the terminal leaflet when it was damaged or difficult to
clamp into the gas exchange chamber. When a leaflet was damanged during
gas exchange measurements, we collected anatomical data from the nearest
leaflet on the same leaf (10.0)\%. We measured chlorophyll concentration
index (CCI) using a chrolophyll concentration meter (MC-100, Apogee
Instruments, Logan, Utah, USA) on the lamina of focal leaflets before
gas exchange measurements at the same time we measured leaf temperature.

\hypertarget{amphistomy-advantage}{%
\paragraph{Amphistomy advantage}\label{amphistomy-advantage}}

We estimated `amphistomy advantage' (\(\mathrm{AA}\)) \emph{sensu}
@parkhurst\_adaptive\_1978, but with modifications previously described
in @triplett\_amphistomy\_2024. \(\mathrm{AA}\) is calculated as the
log-response ratio of \(A\) compared at the same total \(g_\text{sw}\):

@@

We measured the photosynthetic rate of an untreated amphistomatous leaf
(\(A_{\mathrm{amphi}}\)) over a range of \(g_\text{sw}\) values. We
refer to this as an \(A \textendash g_\text{sw}\) curve. We compared the
\(A \textendash g_\text{sw}\) curve of the untreated leaf to the
photosynthetic rate of pseudohypostomatous leaf (\(A_\mathrm{hypo}\)),
which is the same leaf but with gas exchange through the upper surface
blocked by a neutral density plastic (propafilm).

We measured \(A \textendash g_\text{sw}\) curves using a portable
infrared gas analyzer (LI-6800PF, LI-COR Biosciences, Lincoln, Nebraska,
USA). Light-acclimated plants were placed under LEDs dimmed to match
their light treatment during gas exchange measurements. We estimated the
photosynthetic rate (\(A\)) and stomatal conductance to water vapor
(\(g_\text{sw}\)) at ambient CO\(_2\) (415 ppm) and
\(T_\mathrm{leaf} = \qty{25.0}{\degreeCelsius}\). The irradiance of the
light source in the pseudohypo leaf was higher because the propafilm
reduces transmission. To compensate for reduced transmission, we
increased incident \(\mathrm{PPFD}\) for pseudohypo leaves by a factor
1/0.91, the inverse of the measured transmissivity of the propafilm. We
also set the stomatal conductance ratio, for purposes of calculating
boundary layer conductance, to 0 for pseudohypo leaves following
manufacturer directions.

We collected four \(A \textendash g_\text{sc}\) curves per leaf, an
amphi (untreated) curve and a pseudohypo (treated) curve at high
light-intensity
(\(\mathrm{PPFD} = \qty{2000}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\))
and low light-intensity
(\(\mathrm{PPFD} = \qty{150}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\)).
We always measured high light-intensity curves first because
photosynthetic downregulation is faster than upregulation in these
species. To control for order effects, we alternated between starting
with amphi or pseudohypo leaf measurements. Unlike
@triplett\_amphistomy\_2024, preliminary experiments with \emph{Solanum}
indicated a strong order effect in that \(A\) declined in the second
curve. Therefore, we made measurements over two days. On the first day,
we measured high and low light-intensity curves for either amphi or
pseudohypo leaves; on the second day, we measured high and low
light-intensity curves on the other leaf type.

In all cases, we acclimated the focal leaf to high light
(\(\mathrm{PPFD} = \qty{2000}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\))
and high relative humidity (\(\mathrm{RH} = 70\%\)) until \(A\) and
\(g_\text{sw}\) reach their maximum. After that, we decreased
\(\mathrm{RH}\) to \(\approx 10\%\) to induce rapid stomatal closure
without biochemical downregulation. Hence, \(A_\text{amphi}\) and
\(A_\text{hypo}\) were both measured at low chamber humidity after the
leaf had acclimated to high humidity. All other environmental conditions
in the leaf chamber remained the same. We logged data until
\(g_\text{sc}\) reached its nadir. We then acclimated the leaf to low
light
(\(\mathrm{PPFD} = \qty{150}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\))
and 70\% \(\mathrm{RH}\) before inducing stomatal closure with low
\(\mathrm{RH}\) and logging data as described above.

\hypertarget{light-response-curves}{%
\paragraph{Light-response curves}\label{light-response-curves}}

We measured light-response curves on the same leaflets as
\(A \textendash g_\text{sc}\)

testing \gsc

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\addtokomafont{disposition}{\rmfamily}
\usepackage{siunitx}
\newcommand{\aax}{$\mathrm{AA}$}
\newcommand{\Aamphi}{$A_{\mathrm{amphi}}$}
\newcommand{\Ahypo}{$A_{\mathrm{hypo}}$}
\newcommand{\agcurve}{$A \textendash g_\text{sw}$}
\newcommand{\caequals}[1]{$C_\mathrm{a} = \qty{#1}{\micro\mol\raiseto{-1}\mol}$}
\newcommand{\gsw}{$g_\text{sw}$}
\newcommand{\ppfd}{$\mathrm{PPFD}$}
\newcommand{\ppfdequals}[1]{$\mathrm{PPFD} = \qty{#1}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}$}
\newcommand{\rh}{$\mathrm{RH}$}
\newcommand{\rhequals}[1]{$\mathrm{RH} = #1\%$}
\newcommand{\tleafequals}[1]{$T_\mathrm{leaf} = \qty{#1}{\degreeCelsius}$}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Unknown},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle

\subsection{Notes}\label{notes}

@@

\subsection{Main}\label{main}

@@

We distinguished among these hypotheses by comparing AA among wild
tomato species from different native light habitats, grown under
simulated sun and shade light treatments, and measured under contrasting
light intensity (Figure of hypotheses and predictions). We measured AA
on 600 individual plants from 30 accessions (average of 10 replicates
per light treatment) using a recently developed method (\emph{1}). With
this method, we directly compare the photosynthetic rate of an untreated
amphistomatous leaf to that of the same leaf with gas exchange blocked
through the adaxial (upper) surface by transparent plastic, which we
refer to as `pseudohypostomy'. To compare amphi- and pseudohypostomatous
leaves at identical whole-leaf \(g_\text{sc}\), we measure \(A\) over a
range of \(g_\text{sc}\), inducing stomatal opening and closure by
modulating humidity (see Materials and Methods for further details).

@@

\subsection{Materials and Methods}\label{materials-and-methods}

@@

\subsubsection{Accessions}\label{accessions}

We compared AA among 30 ecologically diverse accessions of wild tomato,
including representatives of all described species of \emph{Solanum}
sect. \emph{Lycopersicon} and sect. \emph{Lycopersicoides} (\emph{2})
and the cultivated tomato \emph{S. lycopersicum} var.
\emph{lycopersicum} (tab:accessions). Due to constraints on growth space
and time, we spread out measurements over 80 weeks from May 1, 2022 to
October 31, 2023. Replicates within accession were evenly spread out
over this period to prevent confounding of temporal variation in growth
conditions with accession. {[}anything else to say here? maybe explain
accession selection and phylogeny?{]}

@@

\caption{\label{tbl-accessions}Solanum accessions}

\centering{

}

\end{table}%

\subsubsection{Plant growth conditions}\label{plant-growth-conditions}

@@

\paragraph{Germination and seedling
stage}\label{germination-and-seedling-stage}

@@

\paragraph{Light treatments}\label{light-treatments}

@@

We assigned sun and shade treatment to lower and upper racks of a
\(\qty{1.22}{\meter} \times \qty{2.44}{\meter}\) shelving unit in a
climate-controlled growth room. We assigned the sun treatment to the
lower rack to limit diffuse light from reaching the shade treatment. The
average daytime \ppfd{} was \ppfdequals{800}
(\(\qty{800}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\)) and
\ppfdequals{125}
(\(\qty{125}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\)) for sun
and shade treatments, respectively. To isolate the effect of light
intensity from quality, we used the same LED model with the the same
spectrum (Fluence SPYDR 2i, Austin, Texas, USAS), but dimmed the lights
in the shade treatment. To maintain homogeneous environmental conditions
other than light, we mixed air within the growth room using an air
circulator (Vornado 693DC, Andover, Kansas, USA) and within racks using
a miniature oscillating air circulator (Vornado Atom 1, Andover, Kansas,
USA). Despite these efforts, the air in the sun treatment was on average
\(\qty{1}{\degreeCelsius}\) warmer and the average RH was consequently 5
lower. However, because of evaporative cooling, the leaves in the sun
treatment were only \(\qty{1}{\degreeCelsius}\) on average (\(n = 600\)
leaves).

\subsubsection{Leaf trait measurements}\label{leaf-trait-measurements}

We selected a fully expanded, unshaded leaf at least six leaves above
the cotyledons during early vegetative growth. This typically meant that
plants had grown in light treatments for \(\approx\) 4 weeks, ensuring
they had time to sense and respond developmentally to the light
intensity of the treatment rather than the seedling conditions
(\emph{3}). Shade plants grew slower than sun plants, hence leaves at
the same developmental stage were measured on chronologically older
plants in the shade treatment. In some sun plants, we had to use leaves
higher on the stem because short internodes made lower leaves
inaccessible with the gas exchange equipment. We measured terminal
leaflets in 70.0\% of cases, but used the lateral leaflet closest to the
terminal leaflet when it was damaged or difficult to clamp into the gas
exchange chamber. When a leaflet was damaged during gas exchange
measurements, we collected anatomical data from the nearest leaflet on
the same leaf (10.0)\%. We measured chlorophyll concentration index
(CCI) using a chrolophyll concentration meter (MC-100, Apogee
Instruments, Logan, Utah, USA) on the lamina of focal leaflets before
gas exchange measurements at the same time we measured leaf temperature.

\paragraph{Amphistomy advantage}\label{amphistomy-advantage}

We estimated `amphistomy advantage' (\aax) \emph{sensu} (\emph{4}), but
with modifications previously described in (\emph{1}). \aax{} is
calculated as the log-response ratio of \(A\) compared at the same total
\gsw:

@@

We measured the photosynthetic rate of an untreated amphistomatous leaf
(\Aamphi) over a range of \gsw{} values. We refer to this as an
\agcurve{} curve. We compared the \agcurve{} curve of the untreated leaf
to the photosynthetic rate of pseudohypostomatous leaf (\Ahypo), which
is the same leaf but with gas exchange through the upper surface blocked
by a neutral density plastic (propafilm).

We measured \agcurve{} curves using a portable infrared gas analyzer
(LI-6800PF, LI-COR Biosciences, Lincoln, Nebraska, USA).
Light-acclimated plants were placed under LEDs dimmed to match their
light treatment during gas exchange measurements. We estimated the
photosynthetic rate (\(A\)) and stomatal conductance to CO\(_2\) (\gsw)
at ambient CO\(_2\) (\caequals{415}) and \tleafequals{25.0}. The
irradiance of the light source in the pseudohypo leaf was higher because
the propafilm reduces transmission. To compensate for reduced
transmission, we increased incident \ppfd for pseudohypo leaves by a
factor 1/0.91, the inverse of the measured transmissivity of the
propafilm. We also set the stomatal conductance ratio, for purposes of
calculating boundary layer conductance, to 0 for pseudohypo leaves
following manufacturer directions.

We collected four \agcurve{} curves per leaf, an amphi (untreated) curve
and a pseudohypo (treated) curve at high light-intensity
(\ppfdequals{2000}; 97.7:2.3 red:blue) and low light-intensity
(\ppfdequals{150}; 87.0:13.0 red:blue). We always measured high
light-intensity curves first because photosynthetic downregulation is
faster than upregulation in these species. To control for order effects,
we alternated between starting with amphi or pseudohypo leaf
measurements. Unlike (\emph{1}), preliminary experiments with
\emph{Solanum} indicated a strong order effect in that \(A\) declined in
the second curve. Therefore, we made measurements over two days. On the
first day, we measured high and low light-intensity curves for either
amphi or pseudohypo leaves; on the second day, we measured high and low
light-intensity curves on the other leaf type.

In all cases, we acclimated the focal leaf to high light
(\ppfdequals{2000}) and high relative humidity (\rhequals{70}) until
\(A\) and \gsw{} reach their maximum. After that, we decreased \rh{} to
\(\approx 10\%\) to induce rapid stomatal closure without biochemical
downregulation. Hence, \Aamphi{} and \Ahypo{} were both measured at low
chamber humidity after the leaf had acclimated to high humidity. All
other environmental conditions in the leaf chamber remained the same. We
logged data until \gsw{} reached its nadir. We then acclimated the leaf
to low light (\ppfdequals{150}) and \rhequals{70} before inducing
stomatal closure with low \rh and logging data as described above.

\paragraph{Light-response curves}\label{light-response-curves}

In 90.0\% of plants, we measured light-response curves on the same
leaflets as \agcurve{} curves. However, when a leaflet was damaged
during \agcurve{} curves, we used the next closest leaflet for
light-response curves. Leaves acclimated to high light-intensity
(\ppfdequals{2000}), ambient CO\(_2\) (\caequals{415}), \rhequals{50},
and \tleafequals{25}. After \(A\) and \gsw{} stabilized, we measured
\(A\) at 20 light-intensity levels between \(0\) and
\(\qty{2000}{\micro\mol\raiseto{-2}\meter\raiseto{-1}\second}\) in
descending order.

\paragraph{Stomatal anatomy}\label{stomatal-anatomy}

We estimated the stomatal density and size on ab- and adaxial leaf
surfaces from all leaves, using guard cell length as a proxy for
stomatal size since it proportional to maximum conductance (\emph{5}).
We made surface impressions of leaf lamina from the same area used for
gas exchange measurements using a-silicone impression material (Zhermack
elite HD+, light body, fast set, Rovigo, Italy). We applied clear nail
polish to make positive replicas of the impression. After nail polish
dried, we mounted replicas on a microscope slide using transparent tape
(\emph{6}). We digitized a portion of each leaf surface replica using a
brightfield microscope (Leica DM2000, Wetzlar, Germany). We counted and
measured guard cell length on all stomata using the FIJI implementation
of ImageJ2 version 2.3.0 (\emph{7}), then divided the count by the
visible leaf area (\(\qty{0.890}{\raiseto{2}\milli\meter}\)) to estimate
stomatal density.

\paragraph{Internal leaf anatomy}\label{internal-leaf-anatomy}

Add information on internal leaf anatomy, if using

\paragraph{Leaf mass per area}\label{leaf-mass-per-area}

Leaf mass per area (LMA) is the dry mass divided by the leaflet area. We
scanned fresh leaflets on a flat bed scanner (Epson V600, Los Alamitos,
California. USA) and measured leaflet area from digital images using the
FIJI implementation of ImageJ2 version 2.3.0 (\emph{7}). We dried leaves
for 72 hours at \(\qty{74}{\degreeCelsius}\) in a food dehydrator
(Cosori CP267-FD, Vesync Co., Anaheim, California, USA) and weighed
using a benchtop analytical balance (Ohaus PR64 Analytical Balance,
Parsippany, New Jersey, USA). In \(5.0\%\) we measured LMA on the
adjacent leaflet because the focal leaflet was damaged or wilted while
making surface impressions and we could not reliably estimate area. LMA
data are missing from \(3.0\%\) of individuals because the area or mass
was not recorded at all or recorded incorrectly.

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-triplett_amphistomy_2024}
\CSLLeftMargin{1. }%
\CSLRightInline{G. Triplett, T. N. Buckley, C. D. Muir,
\href{https://doi.org/10.1002/ajb2.16284}{Amphistomy increases leaf
photosynthesis more in coastal than montane plants of {Hawaiian} ʻilima
( \emph{{Sida} fallax} )}. \emph{American Journal of Botany}
\textbf{111}, e16284 (2024).}

\bibitem[\citeproctext]{ref-peralta_taxonomy_2008}
\CSLLeftMargin{2. }%
\CSLRightInline{I. E. Peralta, D. M. Spooner, S. Knapp, Taxonomy of wild
tomatoes and their relatives (\emph{solanum} sect.
\emph{Lycopersicoides}, sect. \emph{Juglandifolia}, sect.
\emph{Lycopersicon}; {Solanaceae}). \textbf{84} (2008).}

\bibitem[\citeproctext]{ref-schoch_dependence_1980}
\CSLLeftMargin{3. }%
\CSLRightInline{P.-G. Schoch, C. Zinsou, M. Sibi,
\href{https://doi.org/10.1093/jxb/31.5.1211}{Dependence of the stomatal
index on environmental factors during stomatal differentiation in leaves
of \emph{{Vigna} sinensis} {L}.: 1. {Effect} of light intensity}.
\emph{Journal of Experimental Botany} \textbf{31}, 1211--1216 (1980).}

\bibitem[\citeproctext]{ref-parkhurst_adaptive_1978}
\CSLLeftMargin{4. }%
\CSLRightInline{D. F. Parkhurst,
\href{https://doi.org/10.2307/2259142}{The adaptive significance of
stomatal occurrence on one or both surfaces of leaves}. \emph{The
Journal of Ecology} \textbf{66}, 367--383 (1978).}

\bibitem[\citeproctext]{ref-sack_developmental_2016}
\CSLLeftMargin{5. }%
\CSLRightInline{L. Sack, T. N. Buckley,
\href{https://doi.org/10.1104/pp.16.00476}{The developmental basis of
stomatal density and flux}. \emph{Plant Physiology} \textbf{171},
2358--2363 (2016).}

\bibitem[\citeproctext]{ref-mott_amphistomy_1991}
\CSLLeftMargin{6. }%
\CSLRightInline{K. A. Mott, O. Michaelson, Amphistomy as an adaptation
to high light intensity in \emph{{Ambrosia} cordifolia} ({Compositae}).
\emph{American Journal of Botany} \textbf{78}, 76--79 (1991).}

\bibitem[\citeproctext]{ref-schindelin_fiji_2012}
\CSLLeftMargin{7. }%
\CSLRightInline{J. Schindelin, I. Arganda-Carreras, E. Frise, V. Kaynig,
M. Longair, T. Pietzsch, S. Preibisch, C. Rueden, S. Saalfeld, B.
Schmid, J.-Y. Tinevez, D. J. White, V. Hartenstein, K. Eliceiri, P.
Tomancak, A. Cardona, \href{https://doi.org/10.1038/nmeth.2019}{Fiji: An
open-source platform for biological-image analysis}. \emph{Nature
Methods} \textbf{9}, 676--682 (2012).}

\end{CSLReferences}
